The Forecast Trap

### 2022-02-23 manuscript/round3-manuscript.Rmd
:pen:

abstract: |
  Encouraged by decision makers' appetite for future information on topics ranging from elections to pandemics, and enabled by the explosion of data and computational methods, model based forecasts have garnered increasing influence on a breadth of decisions in modern society. Using a classic example from fisheries management, I demonstrate that selecting the model or models that produce the most accurate and precise forecast (measured by statistical scores) can lead to decidedly worse outcomes (measured by real-world objectives). This can create a forecast trap, in which the outcomes such as fish biomass or economic yield decline while the manager becomes increasingly convinced that these actions are consistent with the best models and data available. The forecast trap is not unique to this example, but possible whenever (1) the optimal management policy is not unique to the generative process, and (2) the generative process is not in our candidate set of models.

@@

In this paper, I demonstrate that the model that makes the better prediction (rigorously defined as a strictly proper score, @Gneiting2007) is not necessarily the model that makes the better policy (rigorously defined in terms of utility, e.g. expected net present value, @Clark1990).
I show that our best methods for learning about model structure or parameters by repeatedly comparing forecasts to observations can be counter-productive.
Put another way, the value of information (VOI, as measured by the expected utility given that information minus the utility without it; see @Howard1966; @Katz1987), can actually be negative.
When VOI is negative, the decision-maker may become trapped into accepting mediocre outcomes derived from a model that makes accurate forecasts, even when a less accurate model that would generate better outcomes is available.
In our example, the manager will decide the fishery in question simply has low productivity, because such a model yields better predictions, rather than realizing that the low productivity observed is in fact a consequence of the over-harvesting.
This disconcerting situation can arise whenever two conditions are met: (1) the optimal management policy is not unique to the generative process, and (2) the generative process is not included in candidate set of models.
These conditions do not guarantee the trap will occur, only the circumstances in which it cannot be ruled out entirely.

@@

# From Predictive Models to Decision Policies

How do we translate a model-based forecast into a decision?
It is impossible to discuss outcomes associated with a forecast without first agreeing on this process.
In practice, decision-makers may use a forecast in a wide variety of ways in selecting a course of action, including ways which may run counter to the stated objectives of management [@Ludwig1993].
In principle at least, the field of decision theory provides a formal mechanism for determining the optimal strategy given a model forecast.
For instance, a wide range of ecological conservation and management problems can be expressed as a Markov Decision Process (MDP) problems [@Marescot2013].
Existing computer algorithms such as stochastic dynamic programming (SDP) take a probabilistic model _forecast_ (more precisely, the probability $P(x_{t+1} | x_t, a_t)$ of the system being in state $x_{t+1}$ in the next iteration given that it was previously in state $x_t$ and the manager selected action $a_t$) and the _desired management objective_ (i.e. the maximize the expected biomass of species protected or the expected dollar profit of a fishery [see @Clark1990; @Halpern2013]) as input, and return the _decision policy_ which maximizes that objective [@Marescot2013].
This provides a principled way to associate a decision policy with any given forecast model.

Two features of this approach are worth emphasizing.
First, the resulting decision is derived directly from the forecast model and the desired objective.
The SDP algorithm is a reasonable description of the approach any ideal manager would use -- considering all possible outcomes from all possible sequences of actions and selecting the best sequence.
For complex models this process is too laborious even for a computer, and is often simplified by considering only a selection of predetermined policies (as in Management Strategy Evaluation, MSE, @Punt2016), or scenarios (as in scenario analysis, @Polasky2011).
Such shortcuts are often necessary for complex real-world models, but open additional room for error: the policy we derive from a given forecast may perform poorly not because the model forecast was at fault, but because of those simplifying assumptions about possible policies.
To ensure that the forecast trap is not a result of such assumptions about possible policies, we will consider a problem simple enough to solve directly with SDP.
This leads to the second point: the resulting decision policy is optimal, so long as the forecast model is correct.
In this way, the SDP merely stands in for a mathematically precise way in which forecasts are turned into decisions.
Recognizing that the SDP-derived policy (A) comes directly from the forecast model, and (B) gives the optimal policy for said forecast, seems to suggest that the whatever model makes the better forecast will surely also lead to better outcomes (as measured in terms of whatever utility we have chosen to maximize).
While this intuition is no doubt _often_ accurate, our purpose here is to demonstrate that it is by no means _guaranteed_:
it is also possible for the model which makes the better forecast to lead to worse outcomes.

# Ecological Models

I illustrate this problem using an example from fisheries management.
Fisheries are a significant economic and conservation concern worldwide and their management remains an important debate [e.g. @Worm2006; @Worm2009; @Costello2016].
Moreover, their management has been a proving grounds for theoretical and practical decision-making issues which are widely applicable in other areas of ecology and conservation [@Ludwig1993; @Lande1994], and one that has long wrestled with issues of uncertainty in the context of management decisions [e.g. @Clark1973; @Reed1979; @Walters1981; @Ludwig1982].

While methods such as iterative forecasting [@Dietze2018] and adaptive management [@Walters1978] can be _applied_ to real-world using empirical data, we can only _evaluate_ their potential in hypothetical examples when the true model is known, e.g. through numerical simulation.
That approach allows us to compare both predictions and outcomes across implementations in independent identical replicate worlds.
As noted above, we will assume our underlying model simple enough to solve by SDP, ensuring any poor outcomes from a given forecast are not merely an artifact of an imperfect decision process.
Simple models also have the virtue in being accessible to closed form analysis, which, as we shall see, can give greater insight into when and why this forecast trap arises.
That insight will in turn will allow us to examine if the same problem is likely to arise under more complex models.

<!--
While modern fisheries management frequently relies on complex models which may contain over 100 parameters to reflect the specific age or stage structure of a specific fish stock [@ramlegacy2018; @ram], fisheries management and research has always appreciated the central insights offered by simpler one-dimensional models which continue to underpin both the theory and practice.
For example, the strategies of constant mortality [@Schaefer1954] and constant escapement [@Clark1973; @Reed1979] are built on the same class of models considered here, which also continue to underpin global analyses of stock-rebuilding [e.g. @Costello2016; @Memarzadeh2019] and the foundations of adaptive management [e.g. @Walters1978; @Smith1981; @Walters1981].
While more complex models are often required for specific applications, such models both share features with and are guided by theory built on much more general, simpler models [@Levins1966],
from the Maximum Sustainable Yield of @Schaefer1954 or the epidemiological reproductive number, $R_0$ of @Kermack1927. 
-->

The sustainable harvest decision problem can be stated as follows: The fish stock is observed to be in state $X_t$ at time $t$, and is then subjected to some harvest $H_t$ before recruiting new fish, subject to stochastic environmental noise $\xi_t$, to bring the stock to $X_t+1$,

\begin{equation}
X_{t+1} = f(X_t - H_t, \xi_t) 
\end{equation}

A manager seeks each year to select the harvest quota $H_t$ which will maximize the sum
of the utility derived from such a harvest and such a state, $U(X_t,H_t)$, over all time,
subject to discount rate $\delta$ [@Clark1973]:

\begin{equation}
\sum_{t=0}^{t=\infty} U(X_t, H_t) \delta^t \label{utility}
\end{equation}

We will assume we have been given a fixed price of fish $p=1$ with no additional cost on additional harvest, $U(X_t, H_t) = p \min(H_t, X_t)$ modest discount $\delta = 0.99$.

Let us assume that our set of candidate models are simply the possible parameterizations of a stochastic version of the classic Gordon-Schaefer model [@Schaefer1954; @Gordon1954]:

\begin{equation}
f_i(Y_{t+1}) = Y_t + r_i Y_t \left( 1 - \frac{Y_t}{K_i} \right) * \xi_t(\sigma)
\end{equation}

Where $Y$ is the population size after harvest, $Y_t = X_t-H_t$ and $\xi_t(\sigma)$ represents log-normal random noise with a mean of unity and log-standard-deviation $\sigma_i$.

<!--
This is where most fisheries biologists will probably stop reading.
Modern stock assessment models frequently have over 100 parameters.
However, the goal of exercise understand how a good forecast can lead to a bad decision even when we use these iterative forecasting methods correctly, not predict the safe harvest levels of Pacific Anchovy in 2022.
If any method fails to select the better model in the simplest cases, it does not often do better in more complex ones, even though the problems are harder to detect.
-->

Before this or any other model can generate a forecast, we must first come up with some parameter estimates.
Because the model includes (log-normal) stochastic growth, no amount of data will make any parameter combination impossible, though certain parameter values are more likely than others.
Remember too that parameter estimates may be derived in other ways than than model fitting, especially when parameters are amenable to biological interpretation. 
We will consider the whole range of possible parameters in a moment, but for simplicity, let us begin by focusing in around two of the most interesting regions of that are already included within that larger parameter space of all possible values for $r$, $K$, and $\sigma$.
Let us take "Model 1" as being given by $r_1 = 2$, $K_1 = 16$, $\sigma_1 = 0.05$, "Model 2" by $r_2 = 0.5$, $K_2 = 10$, $\sigma_2 = 0.075$. 
We can imagine our comparison of these two models as a microcosm of the larger comparison between all possible paremeterizations.

Ecologists will rightly scoff at the simplicity of these models -- the real world is much more complicated.
So it is important to bear in mind that these are not models that seek to approximate the stock dynamics of real world fisheries, only to approximate whatever "true model" we are using to drive the simulation.
In recognition of the fact that real world is always more complex than even our best ecological models, we will assume a "true model" for the simulations that is not in the Gordon-Schaefer class (i.e. our candidate models will never contain the true model), but is not so rich that a Gordon-Schaefer curve would seem a hopelessly poor approximation.

For illustrative purposes, we will thus assume the "true" process to be given by Model 3, which is unknown to the decision-maker, but similar enough to at least one of the candidate models might be considered a reasonable approximation:

\begin{equation}
f_3(Y) = Y + r_3 Y^4 \left( 1 - \frac{Y}{K_3} \right)
\end{equation}

with $r_3 = 0.002$, $K_3 = 10$ and $\sigma_3 = 0.05$.

Certainly, the challenge of choosing which model to base a decision policy on in the real world is much harder than this binary choice between two models, and yet it is sufficient to illustrate the trap. 
We will see later why making the models much more complex does not guarantee that the task becomes easier or that the trap may be ruled out.

# Methods for Managing Under Model Uncertainty

I will use this example to illustrate two alternative approaches for iterative learning over model uncertainty: iterative forecasting and adaptive management.
The central difference in the approaches is that iterative forecasting is premised on the ability to score the predictions of alternative models.
Iterative forecasting is silent on the issue of what to do with those scores, this is left up to the decision-maker.
Adaptive management approaches, by contrast, explicitly seek to integrate probabilities over all candidate models to reach a decision.
I consider each in turn.

## Statistical approaches: Forecasting under "Proper" Scoring Rules

Like many decision problems, the task of setting a sustainable harvest quota appears to hinge on having an accurate forecast: if we can predict to what size the fish stock will increase next year, $X_t+1$, and we know the current stock, $X_t$, then we can sustainably harvest $X_{t+1} - X_t$ without decreasing the biomass over the long term.
Selecting a model based on forecast skill is also justifiable on theoretical grounds, since it reduces the risk of overfitting by comparing model predictions to later observations that were not used to estimate the model [@Gneiting2014].

I illustrate the process of model selection by strictly proper scoring rules using two scenarios.
In Scenario A (passive observation) the fish stock is unharvested and allowed to recover towards carrying capacity (as simulated under our "true" model, Model 3) while comparing the observed stock size in each subsequent time step to the distribution predicted under model 1 and model 2 respectively [Fig 1].
The mean, $\mu_t$ and variance, $\sigma_t$ of the forecast are compared against the true observation $x_t$ using a proper scoring rule given by @Gneiting2007,

\begin{equation}
S(x_t|\mu_t,\sigma_t) = -(\mu_t - x_t )^2 / \sigma_t^2 - \log(\sigma_t) \label{proper}
\end{equation}

for each prediction over 100 replicate simulations of 100 time steps each [Fig 1].

```{r figure1, fig.cap = "Forecast performance of each model. Panels A, B: Step ahead predictions of stock size under unfished (A) and fished (B) scenarios. Error bars indicating the 95% confidence intervals around each prediction, while stars denote the observed value in that year. Because the models make different decisions each year in the fished scenario, the observed stock size in year 2, 3, etc under the management of model 1 (blue stars) is different from that under model 2 (red stars). Panels C, D: corresponding distribution of proper scores across all predictions (100 replicates of 100 timesteps). Higher scores are better, confirming that model 2 makes the better forecasts."}
predictions <- read_csv("../data/predictions.csv") %>% mutate(model = as.character(model))

@@

In Scenario B (actively harvest), I have first solved for the optimal management strategy using the forecast-matrices of both model 1 and model 2 [Fig 1b] using SDP [@Marescot2013; code in the Appendix].
Replicate simulations of the stock are harvested at each time step using the optimal quota dictated by either model's forecasts, according to the SDP. 
The resulting stock sizes in the subsequent timestep are scored against the forecast probabilities of each model using Eq \eqref{proper}.
Model 2 unequivocally outperforms model 1 in both scenarios of passive observation and active harvest.

Despite the clearly superior predictive accuracy of model 2 in both scenarios, the outcomes from management under model 2 are substantially worse.
We can assess such outcomes in less abstract terms than forecasting skill, such as economic value (in dollars) or the ecological value (unharvested biomass).
In our simple formulation of the decision problem, the "utility" the manager seeks to maximize is simply the economic value (net present value: the discounted sum of all profits from future harvests, Eq \eqref{utility}) of harvested fish.
This formulation ignores any utility provided by fish that are not harvested, beyond their contribution to future potential harvests.
While it is possible to include such contributions directly in the utility function being optimized [e.g. @Halpern2013], even without doing so, model 1 maintains both a higher unharvested biomass and also leads to higher economic returns throughout [Fig 2].

```{r figure2, message = FALSE, fig.cap = "Ecological and economic performance of each forecast. Harvest quotas derived from model 1 result in a significantly higher fish stock size than under Model 2 (panel A). Economic returns under model 1 are also substantially higher (panel B)"}

@@

In both scenarios, the careful comparison through proper scoring rules has led us to select the worse-performing model.
Crucially, a manager operating under this selection would have little indication that their model was flawed: both future stock sizes and expected harvest yields consistently match model predictions.
Had we been able to include Model 3 in our forecast comparisons, it would equal or outperform the forecasting skill of both model 1 and model 2 (as guaranteed by the theorem of @Gneiting2007), while also matching or out-performing their economic utility (as guaranteed by the theorem of @Reed1979).
In practice, we never have access to the generating model, so it is reasonable to expect model selection to determine the better approximation.
As we see here, the better approximation for forecasting future states does not in fact lead to better outcomes.

One obvious limitation in this comparison is that scenario B treats each model as fixed over the entire course of the management simulation.
In reality, managers will typically re-estimate model parameters after each subsequent observation.
And rather than consider each model/parameter combination in isolation, managers will generate forecasts which reflect the current uncertainty as to which model or parameter values are most likely. 
Re-estimating model parameters results in adjusting those probabilities.
This approach is characterized by adaptive management for sequential decision problems [e.g. @Smith1981], which I employ in the next section.

## Decision-Theoretic Approaches

Any adaptive management strategy updates posterior distributions over model uncertainty [@Punt2016; @Ludwig1982].
Unfortunately, in this case, any such adaptive updating leads to worse outcomes than the equivalent non-adaptive strategy, in which model uncertainty is held fixed.
I illustrate the application of a passive adaptive management strategy to this simple example, following classic examples for parameter [@Ludwig1982] or structural [@Smith1981] model uncertainty.
Passive adaptive management for a simple sequential decision problem is straightforward to implement over a discrete set of states and actions using dynamic programming with iterative updates [@Smith1981, example code in Appendix].
To demonstrate that the behavior is not driven by failure to explore sufficiently, (active adaptive management), I will assign initial probability that model 2 is true at 1%.
After a single iteration of adaptive learning, these probabilities are completely reversed, with the manager deciding that model 2 is almost certainly correct [Fig 3A]. As before, this results in a management practice with much worse ecological and economic outcomes than would have been realized by a manager who stubbornly clung to model 1 without updating, which achieves a net present value that only 31% that expected under management using model 1 alone [Fig 2].

@@

@@

So far we have considered only two alternative combinations of the parameters $r$, $K$ and $\sigma$. 
This simplifies the calculations, because each unique parameter value combination requires a new run of the SDP algorithm to determine the optimal policy from the corresponding forecast.
Increasing the space of possible models to cover a whole plausible range of parameters $r$ and $K$ does little to resolve this problem [Fig 3B]. 
Iterative updates again quickly dismiss the parameter values assumed by model 1, though with more options to choose from, this probability is spread over a range of seemingly plausible candidate models instead of a single alternative model (see Appendix).
While the adaptive management of additional actions and observations slowly narrow this subset of plausible models, decisions based on this uncertainty prevent fish stocks from recovering fully, and realize lower harvests as a result.
Note that learning under either adaptive management approach (using two models or 42), the decision-maker becomes ever-increasingly convinced that they are using the right model or models.
Future stock sizes fall consistently in the range predicted by the model(s), and consistently outside the range predicted by model 1.
Consequently, each iteration the managers are only more firmly convinced that they are maintaining the fish stock near the biomass that supports maximum sustainable yield, when in fact they are sustaining a harvest regime that is preventing recovery of the stock to the much higher productivity regime which would have been achieved under model 1.

<!-- quick summary of what was demonstrated -->
Given this simple decision problem in which one of the two models leads to better ecological and economic outcomes, current approaches invariably choose the wrong one.
Moreover, despite continuing to collect new observations, the decision maker has no way of realizing their mistake.
The manager is trapped into believing whichever model produces the better forecast, even when this results in decidedly worse objective outcomes.
Re-estimating parameters with as new observations accumulate only reinforces the problem [Fig 3A], and
introducing a larger suite of models, such as our wide range of $r$ and $K$ values, does not escape this trap either [Fig 3B].
Other model choice approaches such as goodness-of-fit, information criteria or cross-validation would all prefer model 2 as well.
Only by including the true model in our set of candidates can we be certain that forecast-based methods will converge on optimal outcomes.

<!-- Intuition, Figure 4 -->
The reason for model 1's seemingly contradictory ability to make good decisions but bad forecasts becomes obvious once we compare both curves to that of the underlying model, model 3.
Looking at plots of the growth rate curves for each model [Fig 4A], it is hardly surprising that all model selection approaches prefer the closely overlapping curve of model 2 to the no-where-close curve of model 1 as the better approximation of model 3.
Nevertheless, the decision policy derived from model 1 forecasts is indistinguishable from that based on the true model [Fig 4B], while the policy derived from model 2 forecasts lead to over-harvesting.
Being closest to the true model's forecast skill never guarantees that we are closest to the true model's optimal policy.

@@

Because even complex models frequently have simple control rules or else we are constrained to simple control rules as a practical matter, it is much more likely that any set of candidate models includes parameterizations that reproduce the true optimal policy than that they reproduce the optimal forecast.
When data is initially limited, those simple models that could generate optimal outcomes may have non-negligible probabilities associated with them.
This sets the stage for the forecast trap, adjustments with additional data or comparisons against alternative models that produce more accurate forecasts erodes those probabilities in favor of models which lead to less desirable outcomes.

@@

>>>>

abstract: |
  Encouraged by decision makers' appetite for future information on topics ranging from elections to pandemics, and enabled by the explosion of data and computational methods, model based forecasts have garnered increasing influence on a breadth of decisions in modern society. Using several classic examples from fisheries management, I demonstrate that selecting the model or models that produce the most accurate and precise forecast (measured by statistical scores) can lead to decidedly worse outcomes (measured by real-world objectives). This can create a forecast trap, in which the outcomes such as fish biomass or economic yield decline while the manager becomes increasingly convinced that these actions are consistent with the best models and data available. The forecast trap is not unique to this example, but possible whenever (1) the optimal management policy is not unique to the generative process, and (2) the generative process is not in our candidate set of models.

@@

In this paper, I demonstrate that the model that makes the better prediction (defined as a strictly proper score, @Gneiting2007) is not necessarily the model that makes the better policy (defined in terms of utility, e.g. expected net present value, @Clark1990).
I show that our best methods for learning about model structure or parameters by repeatedly comparing forecasts to observations can be counter-productive.
Put another way, the value of information (VOI, as measured by the expected utility given that information minus the utility without it; see @Howard1966; @Katz1987), can actually be negative.
When VOI is negative, the decision-maker may become trapped into accepting mediocre outcomes derived from a model that makes accurate forecasts, even when a less accurate model that would generate better outcomes is available.
I present two examples of this "forecast trap" and examine how it arises.
This disconcerting situation is possible whenever two conditions are met: (1) the optimal management policy is not unique to the generative process, and (2) the generative process is not included in candidate set of models.
These conditions do not guarantee the trap will occur, only the circumstances in which it cannot be ruled out entirely.

@@

First, I will introduce a motivating example in which we will consider two reasonable process-based models, A and B. 
Model A will produce very accurate forecasts, but lead to much worse outcomes than Model B.
Though I will establish that these accurate forecasts in Model A are not the result of chance or of over-fitting the data, this example may raise more questions than answers. 
To get a better understanding of when the forecast trap arises, I will turn to a simpler ecological model, to which we may apply more sophisticated decision tools of iterative forecasting and adaptive management.
We will see that these approaches do not avoid the forecast trap either. 
No collection of such examples can establish precisely how common the forecast trap may be in real-world applications.
The examples do establish unequivocally that achieving incrementally ever-more-accurate forecasts does _not guarantee_ better decisions.
I conclude by pointing to a range of established and emerging approaches to quantitative decision-making which are not based on forecasts.
As sophisticated forecasting techniques become more common-place in conservation and ecology, the forecast trap is a reminder that we should not forget about these alternatives.

**A note on models and data**

I will use the term "model" to refer to any set of equations or code that can be used to produce a forecast.
This term thus includes not only process-based models, but could also statistical forecasting methods (e.g. ARIMA, @arima), non-parametric approaches such as empirical dynamical modeling [@Ye2015], or machine learning. 
Most such models must first be calibrated to historical data before they can produce a forecast, e.g. by parameter fitting, expert knowledge, or some other means. 
Different choices for those parameters create different forecasts, I will refer to those different parameterizations as different models.
It is of course possible for a decision-maker to consider forecasts coming from multiple structurally different models simultaneously, and potentially assigning different weights to each model.
As more data becomes available, it is possible to update model parameters, or equivalently, update the weights assigned across models.
I will examine such approaches for model ensembles and model updating further on.

In these examples, we will focus on situations in which our 'data' comes from a model simulation rather than empirical sources.
Simulations are simplifications of the real world -- just because a method works in a simulation is no guarantee that it works in reality.
Conversely, if decision methods are not reliable even when applied to simulated cases, we should be even more cautious in how we use them.
Simulations also allow us to consider many replicates and
conduct experiments that would be often impossible or unethical to perform in the real world: for instance, does a given fishery experience better long-term outcomes when managed according to forecasts derived from model 1 or from model 2?

I shall focus on examples involving fisheries management to illustrate principles shared in many ecological systems.
Fisheries are a significant economic and conservation concern worldwide and their management remains an important debate [e.g. @Worm2006; @Worm2009; @Costello2016].
Moreover, their management has been a proving grounds for theoretical and practical decision-making issues which are widely applicable in other areas of ecology and conservation [@Ludwig1993; @Lande1994], and one that has long wrestled with issues of uncertainty in the context of management decisions [e.g. @Clark1973; @Reed1979; @Walters1981; @Ludwig1982].

# A motivating example

To better understand how a model can produce a more accurate forecast and yet still lead to a worse decision, it may be helpful to start with a concrete example.
Figure \ref{fig1} shows both the forecasts and realized management outcomes of two alternative three-species models, "A" and "B" (see Appendix A) in predicting the population dynamics of striped bass (an economically important fishery) and double-crested cormorants (a target species for conservation) which both feed on a population of river herring (whose abundance we assume is not measured). 
Model A accurately forecasts the abundance of both bass and cormorants well into the future, but the optimal management strategy derived from its forecasts leads to steadily declining species abundances and overall disappointing net utility.
Model B produces substantially less accurate forecasts, but Model B's forecasts also lead to better outcomes, seen both in the population abundances and overall net utility.

This management problem is motivated by a real world example of a herring fishery as described in @Brias2021, 
in which our manager seeks to balance multiple objectives of sustaining the cormorant population while maximizing the economic value from harvesting both herring and striped bass.
In the scenario depicted in Figure \ref{fig1}, I have used a the richer five-species model introduced @Brias2021 to drive the underlying dynamics, which includes three competing species of herring that are preyed upon by both the bass and the cormorants. 
Here, the manager seeks to maximize utility given by the weighted sum of the individual objectives [@Brias2021], placing 50% of the weight on the conservation objective and splits the remaining weight evenly over the harvests for predator (bass) and prey (herring) species.
I have assumed a partially observable system - in this case, the manager measures only the abundance of bass and cormorant species, and not of the three herring species.
In this scenario, I have further assumed the manager must choose a fixed fishing effort for herring and for the bass harvest,
I will consider more dynamic decision-processes later, but it is worth noting that in many real world conservation settings policy choices are highly constrained and frequent adjustment of those policies may be costly or impossible.
Likewise, the assumptions of partially observable system and imperfect models are characteristic of ecological decision-making.
Equations and code for all models in this example are presented in Appendix A.

Both models A & B can be seen as alternative attempts to approximate the "true" model (generative process), which in real systems is always unknown and more complex than any model thereof. 
Model A assumes a three-species Ricker model which closely matches the trophic structure of the "true" five-species model, lumping the three competing herring species into a single variable. 
Because herring abundance is not observed directly, the model parameters related to herring growth are less accurate than other parameters.
Model B also lumps herring species together into a single variable, but fails to reflect the trophic relationship between bass and herring.
Model B also oversimplifies the relationship between cormorant and herring population.
This does not make Model B an unreasonable model out-of-hand -- all models contain such simplifications (e.g. our "true" model does not model the trophic relationship between the herring and its food sources or environmental conditions explicitly either).
Both models are consistent with the limited historical data available to them.

# Iterative updating: _Sequential Decision Processes_

While this example demonstrates that the model which provides the better forecast does not necessarily lead to a better decision, it may raise more questions than it answers.
Why does this happen? 
Is this an isolated example or not?
Can this forecast trap be resolved by more sophisticated approaches to model selection and decision-making?
I now consider scenarios involving iterative forecasts and adaptive management: in which the manager monitors outcomes, compares forecasts to observations and updates model estimates.
Such _sequential decision processes_ are not merely iterative versions of single-decision problems, but are much more challenging.
In the opening example, the manager had to choose the harvest policy for each fish species at the start of the scenario and stick with it.
The ability to select new actions in response to new observations turns that decision into a game of chess: each turn, the manager must consider not only their next move but all possible series of moves.

How do we translate a model-based forecast into a decision policy?
It is impossible to discuss outcomes associated with a forecast without first agreeing on this process.
In practice, decision-makers may use a forecast in a wide variety of ways in selecting a course of action, including ways which may run counter to the stated objectives of management [@Ludwig1993].
In principle at least, the field of decision theory provides a formal mechanism for determining the optimal strategy given a model forecast.
For instance, a wide range of ecological conservation and management problems can be expressed as a Markov Decision Process (MDP) problems [@Marescot2013].
Existing computer algorithms such as stochastic dynamic programming (SDP) take a probabilistic model _forecast_ (more precisely, the probability $P(x_{t+1} | x_t, a_t)$ of the system being in state $x_{t+1}$ in the next iteration given that it was previously in state $x_t$ and the manager selected action $a_t$) and the _desired management objective_ (i.e. the maximize the expected biomass of species protected or the expected dollar profit of a fishery [see @Clark1990; @Halpern2013]) as input, and return the _decision policy_ which maximizes that objective [@Marescot2013].
This provides a principled way to associate a decision policy with any given forecast model.

Two features of this approach are worth emphasizing.
As before, the resulting decision is derived directly from the forecast model and the desired objective.
The SDP algorithm is a reasonable description of the approach any ideal manager would use -- considering all possible outcomes from all possible sequences of actions and selecting the best sequence.
For complex models this process is too laborious even for a computer, and is often simplified by considering only a selection of predetermined policies (as in Management Strategy Evaluation, MSE, @Punt2016), or scenarios (as in scenario analysis, @Polasky2011).
Such shortcuts are often necessary for complex real-world models, but open additional room for error: the policy we derive from a given forecast may perform poorly not because the model forecast was at fault, but because of those simplifying assumptions about possible policies.
To ensure that the forecast trap is not a result of such assumptions about possible policies, we will consider a problem simple enough to solve directly with SDP.
The resulting decision policy is optimal, so long as the forecast model is correct.
In this way, the SDP merely stands in for a mathematically precise way in which forecasts are turned into decisions.
Recognizing that the SDP-derived policy (A) comes directly from the forecast model, and (B) gives the optimal policy for said forecast, seems to suggest that the whatever model makes the better forecast will surely also lead to better outcomes (as measured in terms of whatever utility we have chosen to maximize).
While this intuition is no doubt _often_ accurate, our purpose here is to demonstrate that it is by no means _guaranteed_:
it is also possible for the model which makes the better forecast to lead to worse outcomes.

Let us consider the management of single species in which we seek to maximize the long-term net harvest.
In this scenario, the manager estimates the population size each year and must set the total allowable catch (TAC) for that season.
The underlying dynamics are unknown, but the manager is presented with any of a variety of forecast models which can predict the future stock sizes given the current population size and proposed TAC. 
Our manager does not know which of these models is the most accurate a priori. 
Instead, the manager will be able to compare the population size predicted by each forecast (under the chosen TAC) to the measurement of the population size in the following year before coming up with the next year's catch limit.

Faced with a collection of models, a manager can seek either to identify the best model to use, or to consider an integrative assessment which uses the whole ensemble of models to represent the manager's uncertainty about the underlying process.
I will consider both approaches in turn. 
Figure 2 compares forecasts generated by two of the candidate models to observations drawn from simulations of the underlying process.
As before, these are true forecasts: the model forecasts are generated first, they have not been fit to these observations.
In the un-fished scenario (top panels), both models try to predicting the same un-fished equilibrium dynamics.
In the second scenario (lower panel), the manager uses the optimal SDP policy derived from each forecast to determine the TAC for the following year, and compares the observed stock size to that which the model predicted given that fishing quota.
In both cases, model 2 provides far more accurate forecasts, as seen in the error bars and confirmed by the distribution of proper scores [@Gneiting2007].
Model and simulation details are provided in Appendix B.

```{r figure2, fig.cap = "Forecast performance of each model. Panels A, B: Step ahead predictions of stock size under unfished (A) and fished (B) scenarios. Error bars indicating the 95% confidence intervals around each prediction, while stars denote the observed value in that year. Because the models make different decisions each year in the fished scenario, the observed stock size in year 2, 3, etc under the management of model 1 (blue stars) is different from that under model 2 (red stars). Panels C, D: corresponding distribution of proper scores across all predictions (100 replicates of 100 timesteps). Higher scores are better, confirming that model 2 makes the better forecasts."}
predictions <- read_csv("../data/predictions.csv") %>% mutate(model = as.character(model))

@@

Despite the clearly superior predictive accuracy of model 2 in both scenarios, the outcomes from management under model 2 are substantially worse.
We can assess such outcomes in less abstract terms than forecasting skill, such as economic value them manager sought to optimize (in dollars) or the ecological value (unharvested biomass) [Fig 3].

```{r figure3, message = FALSE, fig.cap = "Ecological and economic performance of each forecast. Harvest quotas derived from model 1 result in a significantly higher fish stock size than under Model 2 (panel A). Economic returns under model 1 are also substantially higher (panel B)"}

@@

A manager operating under model 2 would have little indication that the model was flawed: both future stock sizes and expected harvest yields consistently match model predictions.
This manager would be stuck in the forecast trap, incorrectly mistaking the a degraded ecological state and reduced economic outcomes as the best that can be achieved because future observations continue to validate this model.
Had we been able to include Model 3 in our forecast comparisons, it would equal or outperform the forecasting skill of both model 1 and model 2 (as guaranteed by the theorem of @Gneiting2007), while also matching the economic utility of model 1 (as guaranteed by the theorem of @Reed1979).
In practice, we never have access to the generating model, so it is reasonable to expect model selection to determine the better approximation.
As we see here, the better approximation for forecasting future states does not in fact lead to better outcomes.

## Integrated assessments and adaptive management

Rather than select a single best model (or best parameter value), a manager could choose to integrate over possible outcomes generated from all candidate models.
Updating posterior distributions over parameters and/or weights assigned to different models are examples of this kind of adaptive management [@Punt2016; @Ludwig1982].
I illustrate the application of such an adaptive management strategy, following classic examples for parameter [@Ludwig1982] or structural [@Smith1981] model uncertainty.
I first consider only the same two models considered in the previous example. 
I later consider a larger suite of 42 models, spanning the parameter space of Gordon-Schaefer curves.
To avoid failure to explore sufficiently, [see exploration-exploitation tradeoff, e.g. @Walters1981],
I assign prior belief of 99% weight on the optimally performing model, model 1.  
For comparison, I consider the baseline case in which the manager does not update posterior distribution over which models/parameter values are correct (the manager still chooses a new TAC after each observation, but does not update their belief in the model, i.e. does not _learn_ over time).
The difference between the performance with and without learning is known as the "Value of Information".
In both 2-model and 42-model scenarios, the value of information is strongly negative.
The 2-model case achieves a net present value to -58% of the value of having used model 1 alone [Fig 4]. 
Including all 42 models reduces this to a value of -32%.
Both harvests and fish biomass remain significantly lower under adaptive learning scenarios.

@@

out <- am %>% group_by(method) %>% summarize(total = sum(quota*.99^time))
( out[out$method=="learning","total"][[1]] -  out[out$method=="planning","total"][[1]]) / out[out$method=="planning","total"][[1]]

out <- am_multi %>% group_by(method) %>% summarize(total = sum(quota*.99^time))
( out[out$method=="learning","total"][[1]] -  out[out$method=="planning","total"][[1]]) / out[out$method=="planning","total"][[1]]

@@

<!-- Intuition, Figure 4 -->

The reason for model 1's seemingly contradictory ability to make good decisions but bad forecasts becomes obvious once we compare both curves to that of the underlying model, model 3.
Looking at plots of the growth rate curves for each model [Fig 4A], it is hardly surprising that all model selection approaches prefer the closely overlapping curve of model 2 to the no-where-close curve of model 1 as the better approximation of model 3.
Nevertheless, the decision policy derived from model 1 forecasts is indistinguishable from that based on the true model [Fig 4B], while the policy derived from model 2 forecasts lead to over-harvesting.
Being closest to the true model's forecast skill never guarantees that we are closest to the true model's optimal policy.

@@

@@

# A way forward
